Hot and fresh: evidence of climate-related suboptimal conditions for seagrass in a large Gulf coast estuary

### 2024-02-01 manu-draft.qmd
better characerization of precip trends in text

prp2 <- speidat %>% 
  select(yr, mo, precip_mm) %>%
  filter(mo %in% 6:8) %>% 
  summarise(prptot = sum(precip_mm), .by = yr)
prpmod2 <- lm(prptot ~ yr, prp2)
prpslo2 <- round(coefficients(summary(prpmod2))[2, 1], 2)
prppvl2 <- p_txt(coefficients(summary(prpmod2))[2, 4])
prprsq2 <- round(summary(prpmod2)$adj.r.squared, 2)
prpprd2 <- predict(prpmod2, se = T)
prpstr2 <- paste0(round(prpprd2$fit[1], 1), ' (+/-', round(prpprd2$se.fit[1], 2), ' st. err.)')
prpend2 <- paste0(round(prpprd2$fit[nrow(prp)], 1), ' (+/-', round(prpprd2$se.fit[nrow(prp)], 2), ' st. err.)')
prpdif2 <- round(prpprd2$fit[nrow(prp)] - prpprd2$fit[1], 1)
```

Long-term meteorological data showed increasing trends for air temperature and precipitation ([@fig-meteowqraw]). Mean annual air temperature has increased by `r airtempslo` $^\circ$C per year (`r airtemppvl`, $R^2$ `r airtemprsq`). Mean annual air temperature in 1975 was `r airtempstr` $^\circ$C, whereas current mean annual air temperature in 2022 was `r airtempend` $^\circ$C, showing an overall increase in the period of record of `r airtempdif` $^\circ$C. Similarly, total precipitation during the rainy season has increased by `r prpslo` mm per year, although the model was not significant (`r prppvl`, $R^2$ = `r prprsq`). Removing September from the rainy season showed a significantly increasing trend of `r prpslo2` mm per year (`r prppvl2`, $R^2$ = `r prprsq2`). Using this model, total precipitation in 1975 was `r prpstr2` mm, whereas current total precipitation in 2022 was `r prpend2` mm, showing an overall increase in the period of record of `r prpdif2` mm. The SPI showed notable anomalies in precipitation, with pronounced rainy periods in the early 1980s, late 1990s, 2005, and 2015-2020.

>>>>

prp2 <- speidat %>% 
  select(yr, mo, precip_mm) %>%
  filter(mo %in% 6:8) %>% 
  summarise(prptot = sum(precip_mm), .by = yr)
prpmod2 <- lm(prptot ~ yr, prp2)
prpslo2 <- round(coefficients(summary(prpmod2))[2, 1], 2)
prppvl2 <- p_txt(coefficients(summary(prpmod2))[2, 4])
prprsq2 <- round(summary(prpmod2)$adj.r.squared, 2)
prpprd2 <- predict(prpmod2, se = T)
prpstr2 <- paste0(round(prpprd2$fit[1], 1), ' (+/-', round(prpprd2$se.fit[1], 2), ' st. err.)')
prpend2 <- paste0(round(prpprd2$fit[nrow(prp)], 1), ' (+/-', round(prpprd2$se.fit[nrow(prp)], 2), ' st. err.)')
prpdif2 <- round(prpprd2$fit[nrow(prp)] - prpprd2$fit[1], 1)

seponly <- speidat %>% 
  select(yr, mo, precip_mm) %>%
  filter(mo %in% 9)
sepmod <- lm(precip_mm ~ yr, seponly)
seppvl <- p_txt(coefficients(summary(sepmod))[2, 4])
seprsq <- round(summary(sepmod)$adj.r.squared, 2)
dryonly <- speidat %>% 
  select(yr, mo, precip_mm) %>%
  filter(!mo %in% c(6:9)) %>% 
  summarise(prptot = sum(precip_mm), .by = yr)
drymod <- lm(prptot ~ yr, dryonly)
drypvl <- p_txt(coefficients(summary(drymod))[2, 4])
dryrsq <- round(summary(drymod)$adj.r.squared, 2)
apronly <- speidat %>% 
  select(yr, mo, precip_mm) %>%
  filter(mo %in% 4)
aprmod <- lm(precip_mm ~ yr, apronly)
aprpvl <- p_txt(coefficients(summary(aprmod))[2, 4])
aprrsq <- round(summary(aprmod)$adj.r.squared, 2)
```

Long-term meteorological data showed increasing trends for air temperature and precipitation ([@fig-meteowqraw]). Mean annual air temperature has increased by `r airtempslo` $^\circ$C per year (`r airtemppvl`, $R^2$ `r airtemprsq`). Mean annual air temperature in 1975 was `r airtempstr` $^\circ$C, whereas current mean annual air temperature in 2022 was `r airtempend` $^\circ$C, showing an overall increase in the period of record of `r airtempdif` $^\circ$C. Similarly, total precipitation during the rainy season has increased by `r prpslo` mm per year, although the model was not significant (`r prppvl`). Removing September from the rainy season showed a significantly increasing trend of `r prpslo2` mm per year (`r prppvl2`, $R^2$ = `r prprsq2`; September alone did not have a significant trend, `r seppvl`). Using this model, total precipitation in 1975 was `r prpstr2` mm, whereas current total precipitation in 2022 was `r prpend2` mm, showing an overall increase in the period of record of `r prpdif2` mm. Notably, trends for the dry season (October through May) were not significant (`r drypvl`), except April which had a significantly increasing trend (`r aprpvl`, $R^2$ = `r aprrsq`). The SPI showed notable anomalies in precipitation, with pronounced rainy periods in the early 1980s, late 1990s, 2005, and 2015-2020.
